---
title: "Forecasting US Mortgage Rates"
output:
  html_document: default
  word_document: default
---
We have been gven the data of mortgage rates in US since 2005 to 2019. Using ARIMA the rates have been forecasted for the next 50 weeks.
```{r warning=FALSE, include=FALSE}
library(tseries)
library(urca)
library(forecast)
data<- read.csv('data/MORTGAGE30USsince2005.csv',skip=200)

colnames(data)<-c('year','mortgage')
mortgage<-data$mortgage

```
1)Data Visuualization
```{r}
par(mfrow=c(1,1))
plot.ts(mortgage)
par(mfrow=c(1,2))
acf(mortgage)
pacf(mortgage)
```
From the plot of timeseries, it is evident that the time series is non stationary. Also the ACF decreases gradually which implies that thge time series s non stationary.
Let us perform the ADF test for non-stationarity.

2)Identification of the non-stationarity of the time series
```{r}
adf.test(mortgage)
```
The p-value is 0.43767 which is greater than 0.05. Hence we cannot reject the null that the time series is non stationary.Since it shows a drift, let us check for drift
```{r}
test1<-ur.df(mortgage,type="drift", lags = 8)
summary(test1)

```
The test statistic is not more negative than the critical values, hence we conclude that the time series is non-stationary with a drift

3) Parameter estimation
```{r}
mortgage.diff.one<-diff(mortgage)
adf.test(mortgage.diff.one)

```

After taking the first order difference, we get a p-value of 0.01 which is less than 0.05. Hence we reject the null hypothesis that the time series is non-stationary and conclude that the time series is stationary.
```{r}
par(mfrow=c(1,2))
acf(mortgage.diff.one)
pacf(mortgage.diff.one)

```
The plot shows that the ACF cutts of at 2 and the PACF cutts off at 2. Lets fit an ARMA(2,2) model
```{r}
Arima(mortgage,order=c(2,1,2))
#Checking for other models
Arima(mortgage,order=c(2,1,1))

```
We see that ARIMA(2,1,2) has the lowest AIC, hence let's fit this model.
```{r}
model<-Arima(mortgage,order=c(2,1,2))
```

We see that ar1 is not significant,ma1 is also not significant.
We fit the final model by fixing the ar1 and ma1 to 0.
```{r}
model.revised<-Arima(mortgage,order=c(2,1,2),fixed=c(0,NA,0,NA))
```
Final model:

4) Diagnostic checking
```{r}
checkresiduals(model.revised)
```
We observe that the residuals are random and the ACF is not significant. The p-value is greater than 0.05 hence we cannot reject the null hypotheses that the residuals are white noise

5)Forecast for next 50 weeks
```{r}
forecast(model.revised,h=50)
autoplot(forecast(model.revised,h=50))

```

